
Activity Monitoring
========================================


```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE)

```

## Load library

```{r}
library(ggplot2)
library(dplyr)
``` 

## Reading dataset and process the data


```{r}
setwd("C:/Users/wenju/OneDrive/Documents/R/repoducible_research/week2")
activity <- read.csv("activity.csv",header =TRUE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
str(activity)
```

## Histogram of the total number of steps taken each day

### 1. Calculating the total number of steps taken per day (ignore the missing valuve in the dataset)

```{r, total_day_steps}
day_step<- with(activity, tapply(steps, date, sum, na.rm = TRUE))
day_step
```

### 2. Making a histogram of the total number of steps taken each day

```{r histogram_1}
hist(day_step, main = "Total Daily Steps", xlab= "steps", 
     col = "blue", breaks = 10, xlim = c(0, 25000))
```

## Mean and median of the total number steps taken per day
```{r mean and median}
day_mean <- mean(day_step)
print(paste("mean of the total steps per day: ", day_mean))
day_median <- median(day_step)
print(paste("median of the total steps per day: ", day_median))
```

## Time series plot of the average number of steps taken
### 1. Calculating the average number of steps taken per 5-minute inteval
```{r average_interval_steps}
intv_step<- with(activity, tapply(steps, interval, mean, na.rm = TRUE))
intestep <- data.frame(interval= as.numeric(names(intv_step)),
                       steps = intv_step)
str(intestep)
```

#### 2. Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across 
all days (y-axis)
```{R plot_1}
inteplot<- ggplot(data= intestep, aes(x=interval, y=steps)) +
    geom_line(color= "red") + geom_point(color = "red")
inteplot
```


## Calculateing which 5-minute interval that, on average, contains the maximum number of steps
```{r max_interval}
max_intevel<- filter(intestep, intestep$steps == max(intestep$steps))
maxinterval <-max_intevel$interval
print(paste("The 5-minute interval, on average, conatins the maximum number of steps is: ", maxinterval))
```

## Code to describe and show a strategy for imputing missing data
### 1. Calculating and reporting the total number of missing values in the dataset 
```{r}na_num<-colSums(is.na(activity))
na_num
```

### 2. Devising a strategy for filling in all of the missing values in the dataset. Creating a new dataset in which the missing value are replaced by the the mean for that 5-minute interval.
2.1 method 1
```{r new_dataset}
activity2 <- activity
for (i in 1 :nrow(activity2)){
    if (is.na(activity2$steps[i])) {
        inter<- activity2$interval[i]
        subact<- subset(activity2, interval == inter)
        imean <- mean(subact$steps, na.rm = TRUE)
        activity2$steps[i]<- imean
    }
}
head(activity2)
table(is.na(activity2$steps))
```

2.2 method 2
```{r new_dataset2}
activity3 <- activity %>% group_by(interval) %>%
    mutate(step_mean = mean(steps,na.rm=TRUE))

for (i in 1 :nrow(activity3)){
    if (is.na(activity3$steps[i])){
    activity3$steps[i]<- activity3$step_mean[i]
    }
}
activity3<- activity3%>% select(1:3)
head(activity3)
```

## Histogram of the total number of steps taken each day after missing values are imputed.
### 1. Calculating the total number of steps taken per day after missing values are imputed.
```{r}
day_step2<- with(activity2, tapply(steps, date, sum))
```

### 2. Making a histogram of the total number of steps taken each day after missing values are imputed.
```{r}
hist(day_step2, main = "Total Daily Steps", xlab= "steps", 
     col = "green", breaks = 10, xlim = c(0, 25000))
```
 
### 3. Calculating and reporting the mean and median total number of steps taken per day after missing values are imputed.
day_step2
day_mean2 <- mean(day_step2)
print(paste("mean of the total steps per day after missing values are imputed: ", day_mean2))
day_median2 <- median(day_step2)
print(paste("median of the total steps per day after missing values are imputed: ", day_median2))

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.
### 1. Defining the weekday and weekend and adding a column to indicate it is weekday or wenkend
```{r}
wkds <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
activity2$wDay <- factor((weekdays(activity$date) %in% wkds), 
                   levels=c(FALSE, TRUE), 
                   labels=c('weekend', 'weekday'))
head(activity2)
```

### 2. Calulating the average number of steps taken per 5-minute interval after missing value are imputed, average aross all weekday or weeken days.
```{r}
act2 <- activity2 %>%
    group_by(interval, wDay) %>%
    summarise(stepmean=mean(steps))
head(act2)
```

### 3. Making a Panle plots
```{r plot2}
p<-ggplot(data = act2, aes(interval, stepmean)) +
    geom_line(color = "steelblue", size = 1) +
    geom_point(color = "steelblue") + 
    labs(title = "Average Number of Steps Taken 
         per 5- mintute interval",
         y = "Numeber of steps", x = "Interval") + 
facet_grid(wDay ~ .) 
p
```
